---
title: "Summarizing & Visualizing Data"
subtitle: "The Tidyverse Way"
date: "September 7th, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("tidyverse",dependencies = TRUE)
#library(tidyverse)
library(openintro)
library(dplyr)
library(ggplot2)
data("email")
```

Instructions:  
-   Get into groups of 2-3  
-   You will be randomly assigned a number between 2-21 (that number represents the column of your variable of interest)

### Summarizing

Create summary statistics for your column using `summarise()`. 

```{r}
email %>% 
  summarise(mean = mean(password),
            sd = sd(password),
            count = n())
```

Create summary statistics for your column + spam using `group_by()` 

```{r}
email %>% 
  group_by(spam) %>% 
  summarise(count = n(),
            mean_password = mean(password),
            sd_password = sd(password)
            
                  )
  
```

***

### Visualizing

Create a plot for your column using `ggplot2`. 

```{r}
barplot(table(email$password))

ggplot(data = email, aes(x = password)) +
    geom_bar()

```


Create a plot for your column + spam using `ggplot2`.  

```{r}

email$spam_f <- factor(email$spam,
                       labels = c("No Spam", "Spam"))

email %>% 
  select(spam, spam_f)

ggplot(email, aes(x = spam_f, y = password)) +
    geom_boxplot()
```

