---
title: "Midterm 2"
author: "Ly Nguyen"
date: "2023-11-07"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
#install.packages("gtsummary")
library(dplyr)
library(ggplot2)
library(openintro)
library(knitr)
library(gtsummary)
library(tidyr)
```


```{r}
bcar <- read.csv("~/Documents/Personal Docs_East Bay/STAT 630/bcar.csv")
View(bcar)
```

Part 1: Data Cleaning

```{r}
bcar <- bcar %>%
  mutate(month = factor(month),
         dose = factor(dose),
         sex_f = factor(sex,
                         labels = c("Female", "Male"))
)
summary(bcar)

```

Part 2: Exploratory Data Analysis

2. 

```{r}

#Calculate n(%) of sex

table(bcar$sex_f)

props <- prop.table(table(bcar$sex_f))
props

#Calculate mean (sd) of bmi

mean_bmi <- round(mean(bcar$bmi),2)
mean_bmi

sd_bmi <- round(sd(bcar$bmi),2)
sd_bmi

#Calculate mean (sd) of chol

mean_chol <- round(mean(bcar$chol),2)
mean_chol

sd_chol <- round(sd(bcar$chol),2)
sd_chol

#Calculate n(%) of dose

table(bcar$dose)

props_1 <- round((prop.table(table(bcar$dose))),2)
props_1


#Calculate mean (sd) of age

mean_age <- round(mean(bcar$age),2)
mean_age

sd_age <- round(sd(bcar$age),2)
sd_age

```

| Variable     | mean(sd) or n (%)                              |
|--------------|------------------------------------------------|
|sex           |Female:     36 (`r round((props[1]*100),2)`%)  |
|              |Male:     40 (`r round((props[2]*100),2)`%)    |
|bmi           |mean: `r mean_bmi` (sd: `r sd_bmi`)     |
|chol           |mean: `r mean_chol` (sd: `r sd_chol`)     |
|dose          |0 mg/day:     16 (`r props_1[1]*100`%)  |
|              |15 mg/day:     14 (`r props_1[2]*100`%)  |
|              |30 mg/day:     16 (`r props_1[3]*100`%)  |
|              |45 mg/day:     12 (`r props_1[4]*100`%)  |
|              |60 mg/day:     18 (`r props_1[5]*100`%)  |
|age           |mean: `r mean_age` (sd: `r sd_age`)     |


3. 

```{r}
# calculate the mean and standard deviation of vitamin E (vite) stratified by month
vite_month0 <- bcar %>%
  filter(month == 0) %>%
  select(vite) %>%
  pull()


mean_vite_month0 <- mean(vite_month0)
mean_vite_month0

sd_vite_month0 <- sd(vite_month0)
sd_vite_month0

vite_month11 <- bcar %>%
  filter(month == 11) %>%
  select(vite) %>%
  pull()

mean_vite_month11 <- mean(vite_month11)
mean_vite_month11

sd_vite_month11 <- sd(vite_month11)
sd_vite_month11

# calculate the mean and standard deviation of beta-carotene (bcarot) stratified by month
bcarot_month0 <- bcar %>%
  filter(month == 0) %>%
  select(bcarot) %>%
  pull()


mean_bcarot_month0 <- mean(bcarot_month0)
mean_bcarot_month0

sd_bcarot_month0 <- sd(bcarot_month0)
sd_bcarot_month0

bcarot_month11 <- bcar %>%
  filter(month == 11) %>%
  select(bcarot) %>%
  pull()

mean_bcarot_month11 <- mean(bcarot_month11)
mean_bcarot_month11

sd_bcarot_month11 <- sd(bcarot_month11)
sd_bcarot_month11
```
| Variable     | Beginning of the study                                 | end of the study         |
|--------------|------------------------------------------------|-------------------------|
|              | mean(sd)                              |  mean(sd)      |
|Vite        |mean: `r mean_vite_month0` (sd: `r sd_vite_month0`)     |mean: `r mean_vite_month11` (sd: `r sd_vite_month11`)  |
|bcarot        |mean: `r mean_bcarot_month0` (sd: `r sd_bcarot_month0`)     |mean: `r mean_bcarot_month11` (sd: `r sd_bcarot_month11`)  |

4. 
 
The mean of vitamin E at the beginning of the study was more than that one at the ending. The sd of the vitamin E was quite the same in both of two period, so the variability stays unchanged. The average bcarot at the beginning was just 240, but at the ending it significantly came up to 1416. The sd of bcarot also went higher, so it shows that the variability of bcarot expanded more. 

6. Compare the shape, center, and spread of both histograms in 2-3 complete sentences. Use the statistics
you calculated in Question 3 to add to your comparison.

7. Using a plot of your choice, visualize the relationship between vitamin E (vite) and dose (dose).

```{r}
ggplot(bcar, aes(x = dose, y = vite)) +
  geom_point() +
  labs(title = "Relationship between  vitamin E (vite) and dose (dose)",
       y = "Beta-carotene dose", x = "Vitamine E level")
```
Part 2: Data Analysis

```{r}
bcar_wide <- bcar %>%
tidyr::pivot_wider(names_from = month,
values_from = c(bcarot, vite))
```


8. Create a 95% confidence interval for the true average difference in vitamin E level in month 11 (vite_11) minus month 0 (vite_0). Check that the necessary conditions are satisfied, compute the interval in R, and then interpret the interval in the context of the problem.
Next, we are going to categorize vitamin E into low or high.
```{r}
bcar_wide$vite_low_0 <- ifelse(bcar_wide$vite_0 < 6, "low vit E", "high Vit E")

bcar_wide$vite_low_11 <- ifelse(bcar_wide$vite_11 < 6, "low vit E", "high Vit E")

bcar_wide <- bcar_wide %>%
mutate_if(is.character, as.factor)

bcar_wide <- bcar_wide %>%
  mutate(vite_diff = vite_11 - vite_0)
 
mean_vite_diff <- mean(bcar_wide$vite_diff)
sd_vite_diff <- sd(bcar_wide$vite_diff)

# Calculate standard error
n <- length(bcar_wide$vite_diff)
se_vite_diff <- bcar_wide$sd_vite_diff / sqrt(n)

# Set the confidence level
conf_level <- 0.95

# Calculate the critical t-value
t_stat <- qt((1 + 0.95) / 2, df = n - 1)

p_val <- pt(t_stat, df = n - 1, lower.tail = FALSE)
p_val

t_stat + c(-1, 1) * qt(0.975, df = n - 1) * se_vite_diff 

stat <- mean(mem$Diff)
se <- sd(mem$Diff) / sqrt(nd)

t_stat <- stat / se
t_stat

```

A confidence interval (CI) of 2.026 means that we are 95% confident that the true the true average difference in vitamin E level in month 11 (vite_11) minus month 0 (vite_0) falls within the range of -2.026 and 2.026 .

Self - assessment: M (I try my best in the rush period of time, I am 90% confident that my result makes sense)
