---
title: "UN Data"
date: "November 9th, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(alr4)
library(dplyr)
library(ggplot2)

# Load data
data("UN11")

summary(UN11)
```

**Goal**: Model the relationship between per capita gross domestic product (`ppgdp`) and life expectancy of females (`lifeExpF`).

**Population Model**: 

$Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i$, $\varepsilon_i \overset{iid}{\sim} N(0, \sigma)$

Life expectancy (of females) = $\beta_0$ + $\beta_1$ ppgdp + $\varepsilon$


**Confounders**: (associated with both X and Y)  
-

**Precision Variables**: (associated with Y, but not X)  
- 

**Hypotheses**:

$H_0:$ 

$H_A:$


**Check Conditions**:

Linearity
```{r}
ggplot(UN11, aes(x = ppgdp, y = lifeExpF)) +
  geom_point()
```

Independence: 

Normality of the residuals
```{r}
# Run SLR model
mod <- lm(lifeExpF ~ ppgdp, data = UN11)

# Check Normality Assumption
qqnorm(resid(mod))
qqline(resid(mod), col = "red")

hist(resid(mod))
```

Equal (constant) Variance of the residuals
```{r}
plot(resid(mod) ~ fitted(mod), main = "Residuals vs. Fitted")  
abline(h = 0, col = "red") 
```


```{r}
hist((UN11$lifeExpF))
```


### Transform X

```{r}
UN11 <- UN11 %>% 
  mutate(log_ppgdp = log(ppgdp))
```

Check Assumptions
```{r}
ggplot(UN11, aes(x = log_ppgdp, y = lifeExpF)) +
  geom_point()

# Run model with transformed x
mod_new <- lm(lifeExpF ~ log_ppgdp, data = UN11)

# Check Normality Assumption
qqnorm(resid(mod_new))
qqline(resid(mod_new), col = "red")

# Check Equal (Constant) Variance Assumption
plot(resid(mod_new) ~ fitted(mod_new), main = "Residuals vs. Fitted") 
abline(h = 0, col = "red") 

```


Summary Statistics:

```{r}
UN11 %>% 
  summarise(mean(log_ppgdp), sd(log_ppgdp),
            mean(ppgdp), sd(ppgdp),
            mean(lifeExpF), sd(lifeExpF))
```



