---
title: "Hypothesis Testing for a Difference in Two Proportions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Does the temperature you drink your tea (or coffee) increase your risk of esophageal cancer?

The "[Golestan Cohort Study](https://onlinelibrary.wiley.com/doi/epdf/10.1002/ijc.32220), a population-based propective sutdy of 50,045 individuals aged 40-75 years", followed individuals for roughly 10 years to see if there was an association between the temperature at which a person drinks tea/coffee and their risk of esophogeal cancer.

After some exclusion criteria were set in place, there were **49,996 total participants considered in the study**. (Approximately, ) **19,448 participants drank their tea at temperatures of < 60°C (140°F)** and the remaining **30,548 drank their tea at higher temperatures**. Of those who prefer lower temperatures, **92 developed esophogeal cancer** compared to the **222 who developed cancer in the higher temperature group*. 

Is there significant evidence that those who prefer their tea hotter are more likely to develop esophogeal cancer?



### 1. Write the hypotheses.
* 1 = higher
* 2 = lower

$H_0: p_1 - p_2 = 0 $ 

$H_A: p_1 -p_2 > 0$


### 2. Check conditions.

1. Independence within and between:

*Within: both sample sizes < 10% population of coffee/tea drinkers 
*Between: coffee/tea heat preferences reasonably independent.

2. success/ failure counts

```{r, echo = FALSE}
n1 <- 30548

n2 <- 19448

p1_hat <- 222/n1

p2_hat <- 92/n2
```

All count $\leq 5$
```{r}
n1*p1_hat; n1*(1-p1_hat)

n2*p2_hat; n2*(1-p2_hat)
```

### 3. Calculate test statistic.
```{r}
stat <- p1_hat - p2_hat
null_val <- 0

se <- sqrt(p1_hat*(1-p1_hat)/n1 + p2_hat*(1-p2_hat)/n2)

z_stat <- (stat - null_val)/se
z_stat
```


### 4. Calculate p-value.
```{r}
p_val <- pnorm(z_stat, lower.tail = FALSE)
p_val
```


### 5. Make a decision and conclude in the context of the problem.

Decision: Reject $H_0$

Conclusion: We have enough evidence that the difference in proportion of esophogeal cancer among those in the higher temp group minus those in the lower temp group is greater than 0. 

