---
title: "Chi-Squared Tests"
date: 'October 24th, 2023'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

### The `prop.test()` function

**Ice cream preferences.** Is the percentage of Creamery customers who prefer chocolate ice cream over vanilla less than 80\%? In a sample of 50 customers 60\% preferred chocolate over vanilla.

$H_0$: $p = 0.8$
$H_A$: $p < 0.8$

```{r}
?prop.test
prop.test(x = 30, n = 50, p = 0.8, alternative = "less")
```


## The Chi-Squared Distribution
```{r}
x <- seq(0, 20, 0.1)
chi1 <- dchisq(x, df = 3)
chi5 <- dchisq(x, df = 5)
chi10 <- dchisq(x, df = 10)

plot(x, chi1, type = "l", 
     ylim = c(0, 0.3),
     ylab = "f(x)",
     main = "Chi-Squared Distribution")
lines(x, chi5, lty = 2)
lines(x, chi10, lty = 3)
```

## Chi-Squared Test for Goodness of Fit

Consider a standard package of milk chocolate M&Ms. There are six different colors: red, orange, yellow, green, blue and brown. Suppose that we are curious about the distribution of these colors and ask, do all six colors occur in equal proportion?

Suppose that we have a simple random sample of 600 M&M candies with the following distribution:

212 of the candies are blue.
147 of the candies are orange.
103 of the candies are green.
50 of the candies are red.
46 of the candies are yellow.
42 of the candies are brown.

Option 1: By Hand
```{r}
obs_mm <- c(212, 147, 103, 50, 46, 42)
exp_mm <- rep(100,6)
num_mm_colors <- length(obs_mm)
  
t_stat_mm <- sum( (obs_mm - exp_mm)^2 / exp_mm )

p_val_mm <- pchisq(t_stat_mm, 
                   df = num_mm_colors - 1,
                   lower.tail = FALSE)
p_val_mm <- 1 - pchisq(t_stat_mm, 
                   df = num_mm_colors - 1)
p_val_mm
```

Conclusion: We have enough evidence that proportions of each color of M\&M is not equal, i.e. 1/6.

Option 2: Using R Functions
```{r}
?chisq.test
chisq.test(obs_mm)
```

Now, suppose I have reason to believe that 40\% are blue, 20\% are orange, 20\% are green, 10\% are red, 5\% are yellow, and 5\% are brown. Do we have evidence that the proportion of M\&M colors comes from this distribution?


```{r}
null_prop <- c(0.4, 0.2, 0.2, 0.1, 0.05, 0.05)
exp_mm2 <- null_prop * 600
exp_mm2

t_stat2 <- sum( (obs_mm - exp_mm2)^2 / exp_mm2 )
t_stat2

p_val2 <- pchisq(t_stat2, 
                 df = num_mm_colors - 1,
                 lower.tail = FALSE)
p_val2

test <- chisq.test(obs_mm, p = null_prop)
test
test$expected
```

## Chi-Squared Test for Independence

Suppose 395 people are randomly selected, and are "cross-classified" into one of eight cells, depending into which age category they fall and whether or not they support legalizing marijuana.

  
| | 18-24 | 25-34 | 35-49 | 50-64 | Total |
|:----|:----:|:----:|:----:|:----:|:----|
| Yes | 60 | 54 | 46 | 41 | 201 |
| No | 40 | 44 | 53 | 57 | 194 |
| Total | 100 | 98 | 99 | 98 | 395

Are marijuana support and age group independent?

```{r}
data <- matrix(c(60, 54, 46, 41,
                  40, 44, 53, 57),
                  nrow = 2,
                  byrow = TRUE)
rownames(data) <- c("Yes", "No")
colnames(data) <- c("18-24", "25-34", "35-49", "50-64")

test <- chisq.test(data)
test$expected

test
```

Decision: 

Conclusion: 

