---
output:
  pdf_document: default
  html_document: default
---
---
title: "Normal Probabilities"
output:
  pdf_document: default
  ---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

-   `dnorm()`: P(X = x) = height of normal distribution

-   `pnorm()`: pnorm(x, mean, sd)

    -   lower.tail = TRUE (default) $\Rightarrow \ \ P(X \leq x) = F(x)$
    -   lower.tail = FALSE $\rightarrow \ \ P(X \geq x)$

-   `qnorm()`: Inverse normal - Find c, such that $P(X \leq c) = p$

-   `rnorm()`: Random samples of size n

1.  Let $Z \sim N(0,1)$. Compute the following probabilities.

    a.  $P(Z < 1.35)$ = `r pnorm(1.35)`  

    $f(x) = (2\pi\sigma^2)^{-1/2}...$

```{r}
pnorm(1.35)
```  
  
  mean = 36, sd = 0.11, Find P(X<35.8)
  
  1st way:
  
  P(X<35.8) = P(Z< (35.8-36)/0.11) = P(Z<-1.82) = pnorm(-1.82) = 0.0344
  
  2nd way:
  
  P(X<35.8) = pnorm(35.8, 36, 0.11) = 0.0344
  
b.    $P(Z > -0.27)$

```{r}
pnorm(-0.27, lower.tail = FALSE)
```  
  
c.    $P(-0.1 < Z < 1.4)$

```{r}
pnorm(1.4) - pnorm(-0.1)
```  

2.  Next, find the value $c$ such that:  

    a.  $P(Z < c) = 0.025$

```{r}
qnorm(0.025)
```
  
b.  $P(X > c) = 0.05$ where $X \sim N(10, 5)$  

```{r}
qnorm(0.05, 10, 5, lower.tail = FALSE)
```

c.  $P(-c < Z < c) = 0.99$.

        $c$ = `r qnorm(0.995)`

```{r}
qnorm(0.995) #P(Z < c) = 0.995 => Find c? 
qnorm(0.005, lower.tail = FALSE) #P(Z > c) find c?
qnorm(0.005) #P(Z < c) = 0.005 

#qnorm(0.995) : P(Z < c) => find or qnorm(1-0.995, lower.tail = FALSE)
#qnorm(1-0.995) : P(-c < Z)

```

3.  Generate a random sample of size $n = 1000$ from a standard normal distribution. Use the `graphics` package to plot the histogram. Make sure your plot is well-labeled!

```{r}
x <- rnorm(1000)
x
hist(x, 
     main = "Distribution of 1000 samples \nfrom N(0,1)",
     xlab = "x",
     ylab = "Frequency")
```

4.  Compare your histogram to at least one other person's in the class. Are they the same or different? Why is this the case?

5.  Use the `graphics` package to plot the pdf of a standard normal distribution.

```{r}
x <- seq(-3.5, 3.5, 0.1)
x
y <- dnorm(x)
y
plot(x, y, type = "l")

```

6.  Plot the following 3 distributions in a single figure. $X \sim N(10, 3)$, $X\sim N(10,5)$, and $X \sim N(10, 7)$.

```{r}
x1 <- seq(-11.5, 31.5, 0.1) # N(10, 3)
y1 <- dnorm(x1, mean = 10, sd = 3)

x2 <- seq(-11.5, 31.5, 0.1) #N(10, 5)
y2 <- dnorm(x2, mean = 10, sd = 5)

x3 <- seq(-11.5, 31.5, 0.1) #N(10, 7)
y3 <- dnorm(x3, mean = 10, sd = 7)

plot(x1, y1, type = "l", xlim = c(-12, 32))
lines(x2, y2, lty = 2, col = "cornflowerblue") #The line type is set to dashed (lty = 2)
lines(x3, y3, lty = 3, col = "red") #The line type is set to a different dash pattern (lty = 3)
```

7.  The following questions refer to the `email` dataset from the `openintro` package.

    a.  Plot the distribution of the variable `num_char` using `ggplot2`.

```{r}
library(openintro)
library(ggplot2)
library(dplyr)

data("email")
View(email)
```

```{r}
email %>% 
  mutate(spam = factor(spam, labels = c("No", "Yes")))  %>% 
  ggplot(aes(x = num_char,
             y = line_breaks,
             col = spam, shape = spam)) +
  geom_point(alpha = 0.3) +
  labs(title = "Relationship between Num. of Characters\n & Line Breaks, by Spam Status",
       x = "Num. Characters (in thousands)",
       y = "Line Breaks",
       col = "Spam",
       shape = "Spam")

# email %>% 
#   mutate(spam = factor(spam, labels = c("No", "Yes")))  %>% 
#   ggplot(aes(x = ....,
#              y =  ....,
#              col = ..., shape = ...)) +
#   geom_point(alpha = 0.3, binwidth = ..., col = ...., fill = ....) +
#   labs(title = "Relationship between Num. of Characters\n & Line Breaks, by Spam Status",
#        x = "Num. Characters (in thousands)",
#        y = "Line Breaks",
#        col = "Spam",
#        shape = "Spam")


```

b.  Create a qq-plot to assess the normality of this variable.

```{r}
# hist(email$num_char, breaks = 15)

mn <- mean(email$num_char)
std_dev <- sd(email$num_char)

x <- seq(mn-3*std_dev, mn+3*std_dev, 0.1)
y <- dnorm(x, mean = mn, sd = std_dev)

hist(email$num_char, breaks = 15)
lines(x, y, col = "red")

qqnorm(log(email$num_char))
qqline(log(email$num_char), col = "red")
```

8.  Simulations. Simulate three random samples from a standard normal random variables: one with a sample size n = 10, n = 30, and n = 100. Assess the normality of each of the samples. What do you notice? Why do you think that is happening?

```{r}
x_1 <- rnorm(10)
x_1
hist(x_1, 
     main = "Distribution of 10 samples \nfrom N(0,1)",
     xlab = "x",
     ylab = "Frequency")

x_2 <- rnorm(30)
x_2
hist(x_2, 
     main = "Distribution of 10 samples \nfrom N(0,1)",
     xlab = "x",
     ylab = "Frequency")

x_3 <- rnorm(100)
x_3
hist(x_3, 
     main = "Distribution of 10 samples \nfrom N(0,1)",
     xlab = "x",
     ylab = "Frequency")

x_4 <- rnorm(40)

hist(x_4,
     main = "Distribution of 40 samples \nfrom N(0,1)",
     xlab = "x",
     ylab = "Frequency")
```
