---
title: "Let's Learn R Markdown"
date: "2023-08-31"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("openintro") ## AHH NEVER DO THIS
library(openintro)
```

# Hello world.
## Hello world.
### Hello world.
#### Hello world.
##### Hello world.

****  

*Hello* **world**.
~~Hello world~~.

> Add two spaces at the end of a line.

Hello world.  
Hello world.

### Creating R chunks

: Keyboard shortcut: command (ctrl) + option + i.

```{r loading_data}
# Hello world.
data("mtcars")
#data("mtcar")
View(mtcars)
```

: R chunk options

Where is it??
```{r chunk_options, eval=FALSE, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
# Hello world.
summary(mtcars)
#summary(mtcars)
```

Here it is!
```{r}
summary(mtcars)
```

## PLEASE DO NOT DO THIS....EVER
```{r}
mtcars
```

```{r}
mean(mtcars$mpg)
# mean(mtcars$mpg)
```

Or you can write inline code.. The mean mpg for cars in 1974 is `r round( mean(mtcars$mpg), 2)`.

```{r}
# mean_mpg <- mean(mtcars$mpg)
# mean_mpg_1 <- round(mean(mtcars$mpg), 2)
# sd_mpg <- sd(mtcars$mpg)

mean_mpg <- mean(mtcars$mpg)
sd_mpg <- round(sd(mtcars$mpg),2)

mean_wt <- round( mean(mtcars$wt), 2)
sd_wt <- round( sd(mtcars$wt), 2)

mtcars <- factor(mtcars$am)

# mtcars$am_f <- factor(mtcars$am,
#                       labels = c("automatic", "manual"))

table(mtcars$am_f)

addmargins(table(mtcars$am_f))

# 
# mtcars$am_f <- factor(mtcars$am,
#                       labels = c("automatic", "manual"))
# 
# table(mtcars$am_f) # WAIT! BE CAREFUL HERE
# 
# summary(mtcars$am_f)
# 
# 
# addmargins(table(mtcars$am_f))
# 
# mean_mpg <- mean(mtcars$mpg)
# 
# mean_mpg_1 <- round(mean(mtcars$mpg), 2)
# 
# sd_mpg <- sd(mtcars$mpg)
# 
# mean_wt <- round(mean(mtcars$wt), 2)
# sd_wt <- round(sd(mtcars$wt), 2)
# 
# mtcars$am_f <- factor(mtcars$am,
#                               labels = c("automatic", "manual"))
```


```{r}
table(mtcars$am_f)
```


```{r}
# table(mtcars$am_f) / nrow(mtcars)

addmargins(table(mtcars$am_f))
table(mtcars$am_f)/ nrow(mtcars)

props <- prop.table(table(mtcars$am_f))
props
# # ## OR
# props <- prop.table(table(mtcars$am_f))
# props

# table(mtcars$am_f) / nrow(mtcars)


```
#Making table by hand

| Right | Left | Default | Center |
|------:|:-----|---------|:------:|
| 12 | 12 | 12 | 12 |
| 123 | 123 | 123 | 123 |
| 1 | 1 | 1 | 1 |

| Variable | mean (sd) or n (%) |
|--------|--------------|
| mpg | `r mean_mpg` (`r sd_mpg`) |
| weight | `r mean_wt` (`r sd_wt`) |
| Transmission | |
| - Automatic | 19 (`r props[1]*100`%) |
| - Manual | 13 (`r props[2]*100`%) |

```{r}
# summarizing lots of variables
apply(mtcars[, 1:3], 2, mean) 
#indicates that you want to apply the mean() function to each column of the selected subset of the mtcars dataset (which contains the first three columns).
# apply(..., 1, ...) would apply the function to each row.
# apply(..., 2, ...) applies the function to each column.

apply(mtcars[,1:5], 2, mean)

apply(mtcars[,1:6], 2, sd)

```
```{r}
# #####################################################
# # STOP HERE... BACK TO SLIDES
# 
# # Scatterplot matrix (plotting the entire dataset)
#plot(mtcars)
# 
# Plotting a quantitative variable

# plot(mtcars$mpg) # hmm... not helpful
# hist(mtcars$mpg) # better, but boring

hist(mtcars$mpg,
     main = "Distribution of MPG in 1974", # Add a title
     breaks = 10, # specify number of bars
     xlab = "MPG", # Change x label
     ylab = "Count", # Change y label
     col = "blue") # Change color of bars

hist(mtcars$disp,
     main = "Distribution of DISP in 1974",
     breaks = 10,
     xlab = "Disp",
     ylab = "Counts",
     col = "green")

hist(mtcars$wt,
     main = "Dist of wt in 1974",
     breaks = 10,
     xlab = "wt",
     ylab = "count",
     col = "yellow")
# # 
```


```{r}
# Plotting a categorical variable
plot(mtcars$am_f,
     main = "1974 Transmission Type",
     ylab = "Frequency")

# plot(mtcars$am_f,
#      main = "1974 Transmission Type",
#      ylab = "Frequency")


# barplot(table(mtcars$am_f),
#         main = "transmission type")

barplot(table(mtcars$am_f),
        main = "1974 Transmission Type")

# plot(mtcars$am_f,
#      main = "1974 Transmission Type",
#      ylab = "Frequency")
#
# # Making a scatterplot
# # plot(x, y)
#
# plot(mtcars$mpg, mtcars$wt) # BORING
plot(mtcars$drat, mtcars$wt,
     main = "Relationship between \ndrat & Weight",
     xlab = "Drat",
     ylab = "Weight (1000 lbs)",
     pch = 2,
     col = "red",
     cex = .5)

plot(mtcars$mpg, mtcars$wt,
     main = "Relationship between\nWeight & MPG",
     xlab = "MPQ",
     ylab = "Weight (1000 lbs)",
     type = "p",
     pch = 2,
     col = "blue",
     cex = .5)

plot(mtcars$mpg, mtcars$wt,
     main = "Relationship between \nWeight and MPG",
     xlab = "MPG",
     ylab = "Weight (1000 lbs",
     type = "p",
     pch = 2,
     col = "green",
     cex = 0.5)
# 
# 
# # Installing packages
# 
# # install.packages("openintro") # Run once and never again!
# library(openintro)
# 
# ?openintro
# 
# data("babies")

```
```{r}
knitr::kable(head(mtcars[, 1:4]), "pipe")

knitr::kable(head(mtcars[, 1:4]), "simple")

d1 <- head(cars, 3)
d2 <- head(mtcars[, 1:3], 5)
knitr::kable(
  list(d1, d2),
  caption = 'Two tables placed side by side.',
  booktabs = TRUE, valign = 't'
)
```

